The variability of Betelgeuse explained by surface convection

### 2024-05-06 main.tex
Revision

In section \ref{Section 2} we describe the dataset of linear polarization obtained with the Telescope Bernard Lyot (TBL) and the Lomb-Scargle technique used to derive periods. 
In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming. 
In section \ref{Conclusion}, we summup our work.

@@

Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot (TBL). These two instruments measure the polarization 
over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
this high sensitivity, the signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the 
spectrum. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only exceptionally do they 
reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
see the linear polarization signal above noise \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. In those 
occasions we have to add up the signals of thousands of lines to reduce the noise and increase the signal-to-noise ratios. This line addition 
is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
linear polarization signal is similar 
in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
of Betelgeuse and other red supergiants as cited in the Introduction.

In this work we shall look into the signals produced through these line-addition techniques, and which produce a pseudo-spectral line in intensity 
and linear polarization which does not belong to any atomic species in particular but to an average of all of them present and emitting in the 
photosphere of Betelgeuse. After such procedures of line addition, these signals carry the coherent signals present in the photospheric lines, but the particularities of this or 
that spectral line are erased. All these data has been presented before by \cite{auriere_discovery_2016}, \cite{mathias_evolution_2018} and 
\cite{lopez_ariste_three-dimensional_2022}, which also give details on observing times and conditions as well as more detailed description 
on the data reduction. \footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
PolarBase (http://polarbase.irap.omp.eu/).}

@@

\subsection{The variability present in the stokes parameters}

Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
we used the Lomb-Scargle periodogram \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, it fits sinusoidal 
using least-squares at each frequency between the first and the last observation. The better the fit, the higher the power attributed to the frequency. 
We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL during 
that span of 10 years.

We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and +60 km/s to span the entire 
signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal towards -20 km/s, 
and interpreted it as  the maximum velocity of the rising plasma, seen from the Earth. The most redshifted signal was found at +40 km/s. 
This velocity is interpreted
as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than +40 km/s can be found 
and can be interpreted as either dark and cold plasma falling back to Betelgeuse, 
or to rising plumes of plasma located in the hidden face of Betelgeuse,
and rising so high that they appear above the limb.

@@

Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profile of respectively 
the intensity, stokes $Q$, stokes $U$ and the linear polarization. 
%The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
For each figure, the uper panel is the Lomb-Scargle periodogram of each velocity bin (black line) and the total average (red line). The green line corresponds to the window function. The lower panel is the Lomb-Scargle periodogram represented for each velocity bin, while the white dotted line refer to 400 and 200 days. The right panel is the mean profile of each stokes parameter for each velocity bin.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Intensity.pdf}
    \caption{Lomb-Scargle periodogram of the LSD profile of intensity.
    The upper panel is the Lomb-Scargle periodogram for each velocity bin (black line) and the average (red line). The green line is the window function.
    In the lower pannel, the two white dashed lines represent respectively the 400 days and the 200 days periods. The right panel is the average intensity profile for each velocity bin.}
    \label{LS intensity}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.pdf}
    \caption{Same as figure \ref{LS intensity} for stokes $Q$. }
    \label{LS Q}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.pdf}
    \caption{Same as figure \ref{LS intensity} for stokes $U$.}
    \label{LS U}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.pdf}
    \caption{Same as figure \ref{LS intensity} for linear polarization: $\sqrt{Q^2+U^2}$.}
    \label{LS linear polarization}
\end{figure}

@@

%\subsection{Intensity}
Figure \ref{LS intensity} shows the Lomb-Scargle periodogram of the LSD of the intensity profile. The 400 and 200 days period seem to be recovered by the periodogram. However, the 400 days period is uncomfortably close to the peak of the window function, while the 200 days period is not. We also notice that the two periods are present at the same HRV. Interestingly, \cite{mathias_evolution_2018} also found this 200 days periodicity with a shorter observation period.

Figure \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodogram of both stokes $Q$ and $U$. It is clear that the periodogram is completely different 
from the intensity one. A strong signal is present in fig. \ref{LS Q} around 200 days. Concerning stokes U in fig. \ref{LS U}, it appears that the main period is around $0.003 \ \mathrm{days^{-1}}$ (330 days). Other period such as the one at 200 days or 250 days ($0.004 \ \mathrm{days^{-1}}$) are present but difficult to trust. From stokes $Q$ and $U$, we recover the 200 days period and the 330 days period is more or less close to the 400 days period found in the literature. 
Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of the total linear polarization of Betelgeuse: $\sqrt{Q^2+U^2}$. 
Interestingly, the periodogram shows a strong power at both 330 days and 200 days, which are the period that we have found in both stokes $Q$ and $U$.

@@

\subsection{The variability of Betelgeuse inferred from polarimetry imaging}

Using linear polarization, \cite{lopez_ariste_convective_2018} have been able to reconstruct images of Betelgeuse that have been successfully compared 
to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
polarization LSD profile using a Marquardt-Levenberg minimisation.

Betelgeuse is observed in average every month by the TBL, and hence
we are able to follow its surface activity through polarimetry imaging. This has  allowed in the past the estimation of  the size and the
 lifetime of the convective cells
at the surface \citep{lopez_ariste_convective_2018}. From these images, we computed a photo-center, which is sensitive to the size and the number of convective cells. 
An homogeneous star will have a photo-center displacement coinciding with barycenter of the star, whereas a star with one or two large convective cells will have a
photo-center displacement more important, up to a few percent of the stellar radius in the case of red supergiants \citep{chiavassa_probing_2022}.

@@

%We computed the Lomb-Scargle periodogram of the displacement of the photo-center.
But, before going any further, we have to warn
 about the interpretation of such images. 
Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our images 
can be 
rotated by $180^\circ$ degree, and the brightness distribution will still fit the observed LSD polarization profiles. 
If a particular observation were to be treated independently of previous observations, the solution found by 
the algorithm could be any of the ambiguous solutions possible. To obtain a continuity 
between the series of images, 
for a given day we used the brightness distribution of the previous day as initial point 
of the  fitting iteration; and for the first image, we start with a 
random brightness distribution as the beginning of the fit. 
Even though we are able to produce a series of images  consistent with each other, it is important 
to keep in mind that the photo-center displacement computed from the series 
will be affected by the choice of the first image. 
To overcome this issue, we decided to compute
the photo-center displacement from 100 different initial parameters and their subsequent series 
of images in an effort to recover from the ensemble
properties that are independent of the choice of the first image.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.pdf}
    \caption{Lomb-Scargle periodogram of the 100 photo-center displacement of Betelgeuse. Each blue lines correspond to a Lomb-Scargle periodogram of one photo-center
     displacement. The black line is the average of the 100 periodogram. The red dashed lines represent the 400 and 200 days period respectively 
     (from left to right). As before, the green line is the window function.}
    \label{LS photocenter}
\end{figure}

Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement of each one of those 100 series (blue lines) and the average Lomb-Scargle periodogram (black line), 
%As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
while the black dashed lines mark the 400 and 
200 days periods respectively. As before, the window function is represented by the green line. Interestingly, we find two bumps around the 400 days period. We cannot directly confirm that we recover the 400 days period from those small bumps, but since polarimetry imaging involves only surface convection, this is an argument that points toward a variability explained by surface convection. 
%First, we find that the Lomb-Scargle power is high around the LSP, though
%uncomfortably close to the auto-correlation peak.
%We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
%concurrence to what was seen in the periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
%But those bumps almost completely disappear after the dimming. Even though it is 
%hard to identify the 400 days period from those bumps alone,
%we still find that there is a difference before and after the dimming. 
%And this is concluded  from images that exclusively involve surface convection.
%The Lomb-Scargle periodogram of the photo-center displacement points towards a variability explained by convection only,
%and this is consistent with what we obtained from linear polarization.
%Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.

\section{The variability of the light curve of Betelgeuse}
\label{section 3}
%Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
%the LSD profiles and the displacement of the photo-center of the inferred photospheric image reconstructions. 
The presence of the same 
periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks in
polarization data  than in the light curve point to a common origin, one that is rather based upon convective dynamics. Since we associate stokes $Q$ and $U$ to proxies of convection, we associate the 200 days period to the typical timescales on which the Stokes parameters evolve. 
The LSD profiles of stokes $Q$ or $U$ will slightly change when we observe Betelgeuse every month, but are completely overhauled 
on time scales of several months. The Lomb-Scargle periodograms appear to capture this spectral dynamic around periods of 200 days.
 Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
on the surface of Betelgeuse.  The fact that this peak is strongly present in stokes $Q$ but less in $U$ points toward a more stochastic origin, which can fully be explained by convection.\

The other period found in the linear polarization and stokes $U$, the 330 days periods, is more or less close to the 400 days period often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days). It has been shown by \cite{lopez_ariste_convective_2018} that large convective cells can live up to one year, thus this period would be associated to the timescale on which the largest granules evolve. This typical timescale has also been observed in the numerical simulations and might play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}. Once again, the 330 days periodicity being more present in stokes $U$ than $Q$ unravels a stochastic process that is due to convective motions, and is consistent with the work of \cite{gray_mass_2008}, who attributed the 400 days period to the large convective cells.  \

Finally, the intensity profile seems to match the 400 days periods, but is uncomfortably close to a peak of the window function of the TBL and a secondary peak close to 200 days. Since we are able to recover the 200 days period in the stokes paramters, we attribute this period to the convective timescales, but concerning the 400 days period, it could be a mix of the typical convective timescales of large granules associated with stochastic pulsations exited by convective cavities. Up to this point, the model developed to interpret linear polarization does not involve pulsations, that is why we interpret the 200 days period as the convective turnover timescale. The 400 days period involve the intensity profile, thus the 400 days period is probably caused by both convective timescales and stochastic pulsations.
 %As showed by
%\cite{lopez_ariste_convective_2018}, some of those convective cells can live up for years. We may speculate
%that the LSP may be related to those long living structures. However the present span of data from the TBL is insufficient to do more than
%just speculate. It should be noticed that  
% the peak is not present in stokes U, and this may point to a more stochastic event than just long-living structures.

@@

After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
10 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
are of our interest in this comparison: the behaviour of the light curve before and after the great dimming.
%and the amplitudes of the peaks 
%at the main cited periods.

@@

The main observation  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
spectra measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
%But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
%amplitude at selected wavelengths but no others. % Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
%observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
%and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
%And such origin cannot be pulsations.
The interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, we associate the 200 days periodicyt to convective dynamics while the 400 days periodicity is associated to both stochastic pulsations and convective timescales of the largest granules. This conclusion is in agreement with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection. 
%the observed periods are related 
%to convective dynamics and not to radial pulsations of the star.

Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the change of variability before and after the great dimming. We speculate that the great dimming event hustled the dynamics of the photosphere, so that the largest granule are destroyed by the turbulent motions caused by the non equilibrium of the photosphere. Since this event, the variability of Betelgeuse has shortened and is in adequacy with the timescale of which evolve the smallest convective cells, that is 200 days. We expect in the future to see the 400 days variability being the dominant period, the time for the photosphere to go back to equilibrium. Longer time 
series of spectropolarimetric observations of Betelgeuse will be required to shed light on this

>>>>

%In section \ref{Section 2} we describe the dataset of linear polarization obtained with the Telescope Bernard Lyot (TBL) and the 
%Lomb-Scargle technique used to derive periods. 
%In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming. 
%In section \ref{Conclusion}, we summup our work.

@@

Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These two instruments measure the polarization 
over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
this high sensitivity, the signal-to-noise ratios per spectral bin are not sufficient to measure the weak polarization signals in individual 
atomic lines. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only exceptionally do they 
reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
see the linear polarization signal above noise \citep{auriere_discovery_2016}. But most commonly, the amplitudes of linear polarization are 
below noise levels and measuring them requires the addition of the signals of thousands of lines to reduce  noise and increase the 
signal-to-noise ratios. This line addition 
is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
linear polarization signal is similar 
in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
of Betelgeuse and other red supergiants as cited in the Introduction.

In this work we shall look into the signals produced through these line-addition techniques, and which produce a pseudo-line in intensity 
and linear polarization which does not belong to any atomic species in particular but to an average of all of them present and emitting in the 
photosphere of Betelgeuse. After such procedures of line addition, these signals carry the coherent signals present in the photospheric lines, but the particularities of this or 
that spectral line are erased. All these data has been presented before by \cite{auriere_discovery_2016}, \cite{mathias_evolution_2018} and 
\cite{lopez_ariste_three-dimensional_2022}, which also give details on observing times and conditions as well as more detailed description 
on the data reduction\footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
PolarBase (http://polarbase.irap.omp.eu/).}.

@@

\subsection{The variability in the Stokes profiles}

Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
we used  Lomb-Scargle periodograms \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, one fits sinusoids to the data
using least-squares, with frequencies sampled between the first and the last observation. The better the fit, the higher the power attributed to a frequency. 
We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL during 
that span of 10 years.

We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and +60 km/s to span the entire 
signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal in linear polarization  towards -20 km/s, 
and interpreted it as  the maximum velocity of the rising plasma, seen from  Earth. The most redshifted signal was often found at +40 km/s. 
This velocity is interpreted
as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than +40 km/s can be found 
and are interpreted as either dark and cold plasma falling back to Betelgeuse, 
or to rising plumes of plasma located in the hidden face of Betelgeuse,
and rising so high that they appear above the limb.

@@

Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profiles of respectively 
the intensity, Stokes $Q$, Stokes $U$ and the total linear polarization. 
%The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
For each figure, the uper panel is the Lomb-Scargle periodogram of each velocity bin (black line) and the total average (red line). The green line corresponds to the window function. 
The lower panel is the Lomb-Scargle periodogram represented for each velocity bin, while the white dotted lines mark the 400 and 200 days periods for 
reference. The right panel shows the mean profile of each Stokes parameter at each velocity bin.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Intensity.pdf}
    \caption{Lomb-Scargle periodogram of the LSD profile of intensity.
    The upper panel is the Lomb-Scargle periodogram for each velocity bin (black line) and the average (red line). The green line is the window function.
    In the lower pannel, the two white dashed lines mark respectively the 400 days and the 200 days periods. The right panel is the average intensity profile for each velocity bin.}
    \label{LS intensity}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.pdf}
    \caption{Same as figure \ref{LS intensity} for Stokes $Q$. }
    \label{LS Q}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.pdf}
    \caption{Same as figure \ref{LS intensity} for Stokes $U$.}
    \label{LS U}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.pdf}
    \caption{Same as figure \ref{LS intensity} for total linear polarization: $\sqrt{Q^2+U^2}$.}
    \label{LS linear polarization}
\end{figure}

@@

%\subsection{Intensity}
Looking into Fig. \ref{LS intensity}, which shows the Lomb-Scargle periodogram of the LSD of the intensity profile, 
we notice that the 400 and 200 days periods seem to be recovered by the periodogram. However, the 400 days period is uncomfortably close to 
the peak of the window function. We also notice that the two periods are present at the same HRV. Interestingly, 
\cite{mathias_evolution_2018} already found this 200 days periodicity in spectroscopic observations, in spite of a shorter observation period.

Figures \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodograms of Stokes $Q$ and $U$. These periodograms are completely different 
from the intensity one. A strong signal is present in fig. \ref{LS Q} around 200 days. Concerning Stokes U, Fig. \ref{LS U}, it appears on the other hand that the main period is around $0.003 \ \mathrm{days^{-1}}$ (330 days). 
Other periods such as those at 200 days or 250 days ($0.004 \ \mathrm{days^{-1}}$) are present but difficult to trust. From both Stokes $Q$ and $U$, we recover the 200 days period, and also a 330 days period which  is more or less close to the 400 days period found in the literature. 
Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of $\sqrt{Q^2+U^2}$, the total linear polarization of Betelgeuse. 
The periodogram recovers the strong powers at both 330 days and 200 days, already found in either Stokes $Q$ and $U$.

@@

\subsection{The variability in polarimetric imaging}

Using linear polarization, \cite{lopez_ariste_convective_2018} were able to reconstruct images of Betelgeuse that have been successfully compared 
to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
polarization LSD profiles, using a Marquardt-Levenberg minimisation.

Betelgeuse is observed in average every month by the TBL, and hence
we are able to follow its surface activity through this technique of polarimetric imaging. This has  allowed in the past the estimation of  
the size and the lifetime of the convective cells
at the surface \citep{lopez_ariste_convective_2018}. From these images, we can also compute a photo-center, which is sensitive to the size and the number of convective cells. 
An homogeneous star will have a photo-center displacement coinciding with the barycenter of the star, whereas a star with one or two large convective cells will have a
photo-center displacement more important, up to a few percent of the stellar radius in the case of \textbf{red giants?} \citep{chiavassa_probing_2022}.

@@

%We computed the Lomb-Scargle periodogram of the displacement of the photo-center.
But, before proceeding to search for periods in the movements of the photo-center, we have to warn about the interpretation of such images. 
Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our 
images can be rotated by $180^\circ$ degrees, and the brightness distribution will still fit the observed LSD polarization profiles. 
If a particular observation were to be treated independently of previous observations, the solution found by 
the algorithm could be any of the ambiguous solutions possible. To obtain a continuity 
between the time series of images, for a given day we use the brightness distribution of the previous day as initial point 
of the  fitting iteration. The first image of the series  starts its minimization iteration with 
random brightness distribution. 
Even though we are able to produce a series of images  consistent with each other, it is important 
to keep in mind that the photo-center displacement computed from the series 
will be affected by that first image of the series. 
To overcome this issue, we decided to compute
the photo-center displacement from 100 different time serieas, each one starting from a different initial image, in an effort to recover 
from such ensemble
properties that are independent of the choice of the first image.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.pdf}
    \caption{Lomb-Scargle periodogram of the 100 series of photo-center displacements of Betelgeuse. Each blue line correspond to a Lomb-Scargle periodogram of one series of photo-center
     displacement. The black line is the average of the 100 periodograms. The red dashed lines mark the 400 and 200 days periods respectively 
     (from left to right). As before, the green line is the window function.}
    \label{LS photocenter}
\end{figure}

Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement of each one of those 100 series (blue lines) and the average Lomb-Scargle periodogram (black line), 
%As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
while the red dashed lines mark the 400 and 
200 days periods respectively. As before, the window function is represented by the green line. We find two bumps around the 400 days period, which are not 
sufficiently significant to confirm the presence of such period on their own. However, they concur with the results of the previous section, 
and since polarimetry imaging involves only surface convection, this is a further argument towards a variability explained by surface convection alone. 
%First, we find that the Lomb-Scargle power is high around the LSP, though
%uncomfortably close to the auto-correlation peak.
%We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
%concurrence to what was seen in the periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
%But those bumps almost completely disappear after the dimming. Even though it is 
%hard to identify the 400 days period from those bumps alone,
%we still find that there is a difference before and after the dimming. 
%And this is concluded  from images that exclusively involve surface convection.
%The Lomb-Scargle periodogram of the photo-center displacement points towards a variability explained by convection only,
%and this is consistent with what we obtained from linear polarization.
%Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.

\section{Explanations for the variability of the light curve of Betelgeuse}
\label{section 3}
%Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
%the LSD profiles and the displacement of the photo-center of the inferred photospheric image reconstructions. 
The presence of the same 
periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks found in
polarization data  compared to those  in the light curve point to a common origin, one that is rather based upon convective dynamics.

\textbf{The 200-days period appears prominently in the Stokes Q and U profiles. We can }
associate the 200-days period to the typical timescales on which the Stokes parameters evolve. 
The LSD profiles of Stokes $Q$ or $U$ will slightly change when we observe Betelgeuse every month, but are completely overhauled 
on time scales of several months. The Lomb-Scargle periodograms appear to capture these dynamics at periods of about 200 days.
 Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
on the surface of Betelgeuse.  The fact that this peak is strongly present in Stokes $Q$ but less in $U$ 
can be interpreted as a temporal situation of how the convective patterns have been distributed in the last 10 years.
%points toward a more stochastic origin, which can fully be explained by convection.\

The other period found in the total linear polarization and Stokes $U$, the 330-days period, is more or less close to the 400 days period 
often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days). It has been found
by \cite{lopez_ariste_convective_2018} that large convective cells can live  one  to two years, and this period could be associated to the 
timescales on which the largest granules evolve. This typical timescale has also been observed in the numerical simulations and might 
play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}. 
Once again we observe this 330-days periodicity to be more present in Stokes $U$ than in $Q$, what may unravel a 
random situation due to convective motions in the last 10 years. This interpretation appears to be  consistent with the work of \cite{gray_mass_2008}, 
who attributed the 400 days period to the large convective cells.

\textbf{Finally, the intensity profiles do  seem to vary in  400 days periods, but the measured peak is uncomfortably close to a 
peak of the window function of the time series observed with the TBL. This 400-days period, it could be a mix of the typical convective 
timescales of large granules associated with stochastic pulsations exited by convective cavities. Up to this point, the model developed to 
interpret linear polarization does not involve pulsations, that is why we interpret the 200 days period as the convective turnover timescale. 
The 400 days period involve the intensity profile, thus the 400 days period is probably caused by both convective timescales and stochastic 
pulsations.}
 %As showed by
%\cite{lopez_ariste_convective_2018}, some of those convective cells can live up for years. We may speculate
%that the LSP may be related to those long living structures. However the present span of data from the TBL is insufficient to do more than
%just speculate. It should be noticed that  
% the peak is not present in stokes U, and this may point to a more stochastic event than just long-living structures.

@@

%After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
%10 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
%are of  interest in this comparison: the behavior of the light curve before and after the great dimming.
%and the amplitudes of the peaks 
%at the main cited per%iods.

@@

The main result  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
spectra, as measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
%But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
%amplitude at selected wavelengths but no others. % Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
%observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
%and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
%And such origin cannot be pulsations.
But the interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, 
we associate the 200-days periodicity to convective dynamics while the 400-days periodicity may be related \textbf{to both stochastic pulsations and convective timescales of the largest granules.} 
This conclusion concurs with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection. 
%the observed periods are related 
%to convective dynamics and not to radial pulsations of the star.

Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the change of variability before and after the 
great dimming. We speculate that the great dimming event hustled the dynamics of the photosphere, and the largest granules were 
destroyed by the turbulent motions caused by the non equilibrium of the photosphere. Since this event, the variability of Betelgeuse has 
quickened  in adequacy with the timescales of  smaller convective cells, about 200-days. Were this be true, we shall expect in the 
coming future to see the 400-days variability re-appear as the dominant period, whenever  the photosphere returns to some equilibrium. 
Longer time  series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.
